openWAR Analysis
========================================================

We can begin working with **openWAR** data with just a few commands:

```{r, message=FALSE}
require(openWAR)
require(mosaic)
```

```{r}
data(MLBAM2013)
ds = makeWAR(MLBAM2013)
players = getWAR(ds)
```

#### The Leaders

Here are our estimates of the top 50 players in baseball.

```{r}
head(players[order(players$RAA, decreasing=TRUE), c("Name", "RAA", "RAA.bat", "RAA.br", "RAA.field", "RAA.pitch")], 50)
```

And the worst 50:

```{r}
head(players[order(players$RAA, decreasing=FALSE), c("Name", "RAA", "RAA.bat", "RAA.br", "RAA.field", "RAA.pitch")], 50)
```

#### Making Comparisons

We have our point estimates, but now let's put some variance estimates on those. Here is a graphical depiction of the composition of David Wright's openWAR:

```{r}
warplot(431151, data=ds)
```

Of course, comparing the openWAR components among players, along with their variance estimates, is most helpful.

```{r}
warplot(c(431151, 502517, 408234, 285078, 518774, 285079), data=ds)
```

There is widespread agreement that the three best players this season have been Miguel Cabrera, Chris Davis, and Mike Trout. We concur, but here is how you might visualize their respective contributions. 

```{r}
warplot(c(408234, 448801, 545361), data=ds)
```